{RILS}-{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-12-01 paper_resources/doc/bmc_jbd/bmc_jbd.tex
Round 2: up to Section 4.1

\usepackage{tikz}
%\usepackage{fullpage}
\usepackage{calc}
\usetikzlibrary{positioning,shadows,arrows,trees,shapes,fit}
\usepackage{blindtext}
\usepackage{pgfplots}
%\pgfplotsset{compat=1.16}
\usepackage{pythonhighlight}
\usepackage{fixme}
\fxsetup{status=draft} % <====== add this line
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{url}
\raggedbottom
%\algnewcommand\algorithmicforeach{\textbf{for each}}
%\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

\subsubsection{Fitness function}\label{sec:fitness}
\fxnote{Marko: Dosao dovde!}

Symbolic regression objective is to determine the expression fitting available data. It is also allowed to obtain some equivalent expression, since there are multiple ways to express some symbolical equation. Although logically sound and intuitive, this objective is not quantifiable during the solution search/training phase, because the goal expression is not known at that point -- only target values for some of the input data are known. Thus, various numerical metrics are being used in literature to guide the symbolic regression search process. The most popular are the coefficient of determination, also known as $R^2$, and the mean squared error (MSE) or root mean squared error (RMSE). Also, the important aspect of the solution quality is the solution expression complexity, which may correspond to the model size of its tree representation. This follows the Occam's razor principle~\cite{costa2020fast}~ that simpler solution  is more likely to be a correct one. 
The search process of \textsc{RILS}-\textsc{ROLS}  is guided by the non-linear combination of $R^2$, RMSE and solution expression size (complexity) presented in Equation~\ref{eq:fitness}.

\begin{equation}
	\label{eq:fitness}
	fit(s) = (2-R^2(s)) \cdot (1+RMSE(s)) \cdot (1+penalty_{size} \cdot size(S))
\end{equation}

@@

\begin{algorithm}
	\hspace*{\algorithmicindent} \textbf{Input}: solution $s$ \\
	\hspace*{\algorithmicindent} \textbf{Output}: local perturbations (1-perturbations) of solution $s$ -- $s_{perturbations}$
	\begin{algorithmic}[1] 
		\Procedure{All1Perturbations}{$s$}
		\State $s_{perturbations} \gets \emptyset$ \label{line:pertInit}
		\State $s \gets$ \texttt{NormalizeConstants}($s$)
		\State $s \gets$ \texttt{Simplify}($s$)
		\State $s_{subtrees} \gets$ \texttt{SubTrees}($s$)
		\For{$n \in s_{subtrees}$} \label{line:forSSStart}
		\State $n_{perturbations} \gets$ \texttt{All1PerturbationsAroundNode}($s$, $n$)
		\State $s_{perturbations} \gets s_{perturbations} \cup n_{perturbations}$	
		\EndFor \label{line:forSSEnd}
		\State \Return $s_{perturbations}$
		\EndProcedure
	\end{algorithmic}
	\caption{Generation of all 1-perturbations of a given solution.}
	\label{alg:pert}
\end{algorithm}

@@

The set of used 1-perturbations is increased in comparison to those used previously. Namely, in addition to those six possible types of node changes, the following four are added:

@@

Change under number 10 is very important -- it performs the general coefficient tuning, unlike OLS that considers only coefficients in linear combinations.

@@

The following 13 algorithms are compared to our approach: \textsc{AI-Feynman}, \textsc{GOMEA}, \textsc{Afp-Fe}, \textsc{Itea}, \textsc{Afp}, \textsc{Dsr}, \textsc{Operon}, the \textsc{Gplearn} from python package \textsc{gplearn}~\cite{stephens2016genetic}, \textsc{Sbp-Gp}, \textsc{Eplex}, \textsc{Bsr}, \textsc{Feat}, \textsc{Ffx}, \textsc{Mrgp}.

For the (13) competitors, the results are exported from \texttt{SRBench} \url{https://cavalab.org/srbench/results/}, reported in~\cite{la2021contemporary}. 
The maximum computation time allowed for each run of \textsc{RILS}-\textsc{ROLS} is set to 1 hour, while the maximal number of fitness function evaluations is set to 1 million. All other algorithms from \texttt{SRBench} used the time limit that was equal or bellow 8 hours runtime and the same fitness evaluation limit of 1 million (see~\cite{la2021contemporary}). We did not use the full time limit of 8 hours because we empirically concluded that there were almost no improvements after 1 hour \textsc{RILS}-\textsc{ROLS} runtime.

@@

\subsection{Datasets and \texttt{SRBench}}

>>>>

\usepackage{tikz}
%\usepackage{fullpage}
\usepackage{calc}
\usetikzlibrary{positioning,shadows,arrows,trees,shapes,fit}
\usepackage{blindtext}
\usepackage{pgfplots}
%\pgfplotsset{compat=1.16}
\usepackage{pythonhighlight}
\usepackage{fixme}
\fxsetup{status=draft} % <====== add this line
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{url}
\raggedbottom
%\algnewcommand\algorithmicforeach{\textbf{for each}}
%\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

\subsubsection{Fitness function}\label{sec:fitness}

Symbolic regression objective is to determine the expression fitting available data. It is also allowed to obtain some equivalent expression, since there are multiple ways to express some symbolical equation. Although logically sound and intuitive, this objective is not quantifiable during the solution search/training phase, because the goal expression is not known at that point -- only target values for some of the input data are known. Thus, various numerical metrics are being used in literature to guide the symbolic regression search process. The most popular are the coefficient of determination, also known as $R^2$, and the mean squared error (MSE) or root mean squared error (RMSE). Also, the important aspect of the solution quality is the solution expression complexity, which may correspond to the model size of its tree representation. This follows the Occam's razor principle~\cite{costa2020fast}~ that simpler solution  is more likely to be a correct one. 
The search process of \textsc{RILS}-\textsc{ROLS}  is guided by the non-linear combination of $R^2$, RMSE and solution expression size (complexity) presented in Equation~(\ref{eq:fitness}).

\begin{equation}
	\label{eq:fitness}
	fit(s) = (2-R^2(s)) \cdot (1+RMSE(s)) \cdot (1+penalty_{size} \cdot size(s))
\end{equation}

@@

\begin{algorithm}
     \textbf{Input}: solution $s$ \\
     \textbf{Output}: local perturbations (1-perturbations) of solution $s$ -- $s_{perturbations}$ \\
	\begin{algorithmic}[1] 
		\Procedure{All1Perturbations}{$s$}
		\State $s_{perturbations} \gets \emptyset$ \label{line:pertInit}
		\State $s \gets$ \texttt{NormalizeConstants}($s$)
		\State $s \gets$ \texttt{Simplify}($s$)
		\State $s_{subtrees} \gets$ \texttt{SubTrees}($s$)
		\For{$n \in s_{subtrees}$} \label{line:forSSStart}
		\State $n_{perturbations} \gets$ \texttt{All1PerturbationsAroundNode}($s$, $n$)
		\State $s_{perturbations} \gets s_{perturbations} \cup n_{perturbations}$	
		\EndFor \label{line:forSSEnd}
		\State \Return $s_{perturbations}$
		\EndProcedure
	\end{algorithmic}
	\caption{Generation of all 1-perturbations of a given solution.}
	\label{alg:pert}
\end{algorithm}

@@

The set of used 1-perturbations is expanded in comparison to those used previously. Namely, in addition to those six possible types of node changes, the following four are added:

@@

Change under point 10 is very important -- it performs the general coefficient tuning, unlike OLS that considers only coefficients in linear combinations.

@@

The following 14 algorithms are compared to our approach: \textsc{AI-Feynman}, \textsc{GOMEA}, \textsc{Afp-Fe}, \textsc{Itea}, \textsc{Afp}, \textsc{Dsr}, \textsc{Operon}, the \textsc{Gplearn} from python package \textsc{gplearn}~\cite{stephens2016genetic}, \textsc{Sbp-Gp}, \textsc{Eplex}, \textsc{Bsr}, \textsc{Feat}, \textsc{Ffx}, \textsc{Mrgp}.

For the (14) competitors, the results are exported from \texttt{SRBench} \url{https://cavalab.org/srbench/results/}, reported in~\cite{la2021contemporary}. 
The maximum computation time allowed for each run of \textsc{RILS}-\textsc{ROLS} is set to 1 hour, while the maximal number of fitness function evaluations is set to 1 million. All other algorithms from \texttt{SRBench} used the time limit that was equal or bellow 8 hours runtime and the same fitness evaluation limit of 1 million (see~\cite{la2021contemporary}). We did not use the full time limit of 8 hours because we empirically concluded that there were almost no improvements after 1 hour \textsc{RILS}-\textsc{ROLS} runtime.

@@

\subsection{Datasets and \texttt{SRBench}}\fxnote{Marko: Dosao dovde!}
